---
title: "Chapter 2 — Notes"
output: pdf_document
---

```{r setup, include=F}
# knitr::opts_chunk$set(echo = FALSE)
rm(list=ls())
```

## 3.1  Sampling from a grid-approximate posterior

* R Code 3.2:
```{r}
n = 1000
p_grid <- seq(from=0, to=1, length.out=n)
prior <- rep(1, n)
likelihood <- dbinom(x=6, size=9, prob=p_grid)
posterior_notnorm <- likelihood * prior
posterior <- posterior_notnorm / sum(posterior_notnorm)
```

Draw 10,000 samples:
* R Code 3.3:
```{r}
samples <- sample(p_grid, prob=posterior, size=n, replace=T)
```

* 3.4:
```{r}
plot(samples)
```

```{r include=F}
library(rethinking)
```

* 3.5:
```{r}
dens(samples)
```

Let's try more samples:

```{r}
par(mfrow=c(2, 2))
dens(sample(p_grid, prob=posterior, size=1e3, replace=T))
dens(sample(p_grid, prob=posterior, size=1e4, replace=T))
dens(sample(p_grid, prob=posterior, size=1e5, replace=T))
dens(sample(p_grid, prob=posterior, size=1e6, replace=T))
```

## 3.2 Sampling to Summarize

### 3.2.1. Intervals of defined boundaries.

The posterior probability that the proportion of water is less than 0.5:

* 3.6:
```{r}
p_grid < 0.5
sum(posterior[p_grid < 0.5])
```

Samples array:
```{r}
head(samples, 100)
```

The same calculation using samples. Add up all samples that lie in the grid < 0.5, and divide by the total number of samples to get the frequency ~ probability:

* 3.7:
```{r}
n = 1e4
samples = sample(p_grid, prob=posterior, size=n, replace=T)
sum(samples < 0.5) / n
```

 How much probability lies between 0.5 and 0.75:
 * 3.8:
```{r}
sample_points = sum(samples > 0.5 & samples < 0.75) 
sample_points
sample_points / n

```

### 3.2.2. Intervals of defined mass.

Boundaries of the lower 80% posterior probability lies:

* 3.9:
```{r}
quantile(samples, probs = .8)
```

Middle 80%, i.e. lying between 10% and 90%:

```{r}
# 3.10
quantile(samples, probs = c(0.1, 0.9))
```

The above are PERCENTILE INTERVALS.

